---
title: "Katherine's Analysis"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(microbiomeMarker)
library(microshades)
library(speedyseq)
library(cowplot)
```

#Define Paths
```{r}
rds <- "/work/kdm65/2021-Dec_merged.rds"
```

Load and check phyloseq object
```{r}
all_mice.ps <- readRDS(rds)
print(all_mice.ps)
```

```{r}
sample_variables(all_mice.ps)
```

Subset for just Katherine's mice and remove bad samples
```{r}
just.katherine.ps <- all_mice.ps %>%
  subset_samples(Owner == "Katherine")
just.katherine.ps
```

```{r}
plot_bar(just.katherine.ps)
```
None of these samples stand out as having super low read counts, so I don't need to prune any out.

#Relative Abundance
```{r}
rel.ps <- just.katherine.ps %>%
  transform_sample_counts(function(x) x/sum(x))
```

#Abundance Plots

Everything together
```{r}
#melt to dataframes
mdf.all <- prep_mdf(rel.ps)

#generate color objects
color_objs.all <- create_color_dfs(mdf.all,
                                   selected_groups =c("Actinobacteriota","Bacteroidota", "Firmicutes", "Verrucomicrobiota"),
                                   cvd = TRUE)
color_objs.all <- reorder_samples_by(color_objs.all$mdf, color_objs.all$cdf)

#Extract
mdf.all <- color_objs.all$mdf
cdf.all <- color_objs.all$cdf
```

```{r}
cb.all.plot <- plot_microshades(mdf.all, cdf.all) +
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(axis.text.x = element_text(size = 8)) +
  theme(strip.text.x = element_text(size = 8))

cb.all.plot
```

Separate mice colonized by BAA-835 and Akk1616
```{r}
BAA835.ps <- rel.ps %>%
  subset_samples(Group == "BAA-835")

BAA835.ps
```

```{r}
Akk1616.ps <- rel.ps %>%
  subset_samples(Group == "Akk1616")

Akk1616.ps
```


```{r}
#melt to dataframes
mdf.BAA <- prep_mdf(BAA835.ps)
mdf.1616 <- prep_mdf(Akk1616.ps)

#generate color objects
color_objs.BAA <- create_color_dfs(mdf.BAA,
                                        selected_groups =c("Actinobacteriota","Bacteroidota", "Firmicutes", "Verrucomicrobiota"),
                                        cvd = TRUE)
color_objs.BAA <- reorder_samples_by(color_objs.BAA$mdf, color_objs.BAA$cdf)

color_objs.1616 <- create_color_dfs(mdf.1616,
                                        selected_groups =c("Actinobacteriota","Bacteroidota", "Firmicutes", "Verrucomicrobiota"),
                                        cvd = TRUE)
color_objs.1616 <- reorder_samples_by(color_objs.1616$mdf, color_objs.1616$cdf)

#Extract
mdf.BAA <- color_objs.BAA$mdf
cdf.BAA <- color_objs.BAA$cdf

mdf.1616 <- color_objs.1616$mdf
cdf.1616 <- color_objs.1616$cdf
```

```{r}
cb.BAA.legend <- custom_legend(mdf.BAA, cdf.BAA, legend_text_size = 15)

cb.BAA.plot <- plot_microshades(mdf.BAA, cdf.BAA) +
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 8)) +
  facet_wrap(~GI_location~Sample_type, scale = "free_x") +
  theme(strip.text.x = element_text(size = 8)) +
  labs(title = "BAA-835")

plot_grid(cb.BAA.plot, cb.BAA.legend, rel_widths = c(1, .25))
```

```{r}
cb.1616.legend <- custom_legend(mdf.1616, cdf.1616, legend_text_size = 15)

cb.1616.plot <- plot_microshades(mdf.1616, cdf.1616) +
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 8)) +
  facet_wrap(~GI_location~Sample_type, scale = "free_x") +
  theme(strip.text.x = element_text(size = 8)) +
  labs(title = "Akk1616")

plot_grid(cb.1616.plot, cb.1616.legend, rel_widths = c(1, .25))
```

```{r}
cb.faceted <- plot_grid(cb.BAA.plot, cb.1616.plot, ncol = 1)
plot_grid(cb.faceted, cb.BAA.legend, rel_widths = c(3,1))
```

Visually, I'm not seeing much of a difference between BAA-835 and Akk1616 colonized mice when looking at each individual sample type/location combination. Comparing colon vs ileum, it looks like Actinobacteriota and Bacteroidota are only present in ileum samples. Comparing lumen vs tissue, there does seem to be relatively more akkermansia in the tissue samples, which is consistent to prior qPCR results.

#Differential Abundance
LEfSe

Filter out low abundance taxa
```{r}
mean_cutoff = 0.001
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

pct0.1_prune.rel <- rel.ps %>%
  filter_taxa(mean_test, prune = TRUE)

genus.ps <- pct0.1_prune.rel %>%
  tax_glom(taxrank = "Genus")
```

Colon vs Ileum
```{r}
location.lef <- lefse(genus.ps,
             class = "GI_location",
             multicls_strat = TRUE)
```

```{r}
plot_ef_bar(location.lef) +
  labs(title = "Colon vs Lumen")
```

Lumen vs Tissue
```{r}
type.lef <- lefse(genus.ps,
             class = "Sample_type",
             multicls_strat = TRUE)
```

```{r}
plot_ef_bar(type.lef) +
  labs(title = "Lumen vs Tissue")
```

BAA-835 vs Akk1616
```{r}
strain.lef <- lefse(genus.ps,
             class = "Group",
             multicls_strat = TRUE)
```

```{r}
plot_ef_bar(strain.lef) +
  labs(title = "Gavage Strains")
```

Colidextribacter and Roseburia (associated w/ BAA-835 colinized mice) have been found to be decreased in people with parkinson's disease https://pubmed.ncbi.nlm.nih.gov/34776854/

Enterorhabdus (associated w/ Akk1616 colonized mice) have been found to be positively associated with indicators of high-fat diet induced obesity https://pubmed.ncbi.nlm.nih.gov/35067837/

#Akkermansia
```{r}
akks.ps <- rel.ps %>%
  subset_taxa(Genus == "Akkermansia") %>%
  tax_glom(taxrank = "Genus")
print(akks.ps)
```

```{r}
akks <- akks.ps %>%
  otu_table() %>%
  as.data.frame()
```