---
title: "Dustin's Analysis"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(microbiomeMarker)
```

#Define Paths
```{r}
rds <- "/work/kdm65/2021-Dec_merged.rds"
```

Load and check phyloseq object
```{r}
all_mice.ps <- readRDS(rds)
print(all_mice.ps)
```

```{r}
sample_variables(all_mice.ps)
```

Subset for just Dustin's mice and remove bad samples
```{r}
just.dustin.ps <- all_mice.ps %>%
  subset_samples(Owner == "Dustin")
just.dustin.ps
```

```{r}
plot_bar(just.dustin.ps)
```
None of these samples stand out as having super low read counts, so I don't need to prune any out.

#Relative Abundance
```{r}
rel.ps <- just.dustin.ps %>%
  transform_sample_counts(function(x) x/sum(x))
```

#Abundance Plots
```{r}
AbundancePlot.SampleName = function(ps, tax_level) {
  ps %>%
    plot_bar(fill = tax_level, x = "SampleName") +
    labs(y = "Relative Abundance") +
    geom_bar(aes_string(color = tax_level, fill = tax_level),
             stat = "identity",
             position = "stack")
}

AbundancePlot.Day = function(ps, tax_level) {
  ps %>%
    plot_bar(fill = tax_level, x = "Day") +
    labs(y = "Relative Abundance") +
    geom_bar(aes_string(color = tax_level, fill = tax_level),
             stat = "identity",
             position = "stack")
}
```

Since there are two day 0 Akk_free samples, faceting by day combines those two samples with a summed relative abundance of 2.0, so I will plot the Akk_free separate from the other groups.

```{r}
akk_free.ps <- rel.ps %>%
  subset_samples(Group == "Akk_free")

AbundancePlot.SampleName(akk_free.ps, "Phylum") +
  facet_grid(~Group, scale="free_x", space="free_x")
```

```{r}
gavaged.ps <- rel.ps %>%
  subset_samples(Group != "Akk_free")

AbundancePlot.Day(gavaged.ps, "Phylum") +
  facet_grid(~Group~Day, , scale="free_x", space="free_x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

It looks like there might be something going on with the actinobacteriota. They're present in the pre-gavage akk free samples and the later PBS samples but are much less prevalent in samples of mice gavaged with akkermansia.

#Pruning
Prune to the top 20 genera
```{r}
rel.ps %>%
  taxa_sums %>%
  sort(decreasing = TRUE) %>%
  names %>%
  head(20) ->
  top20

rel.ps %>%
  prune_taxa(top20, .) ->
  prune20.rel
prune20.rel
```

```{r}
pruned.akk_free.ps <- prune20.rel %>%
  subset_samples(Group == "Akk_free")

AbundancePlot.SampleName(pruned.akk_free.ps, "Genus") +
  facet_grid(~Group, scale="free_x", space="free_x")
```

```{r}
pruned.gavaged.ps <- prune20.rel %>%
  subset_samples(Group != "Akk_free")

AbundancePlot.Day(pruned.gavaged.ps, "Genus") +
  facet_grid(~Group~Day, , scale="free_x", space="free_x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

Trying out a colorblind friendly plotting package
```{r}
library(microshades)
library(speedyseq)
library(cowplot)
```

```{r}
#melt to dataframes
mdf.akk_free <- prep_mdf(akk_free.ps)
mdf.gavaged <- prep_mdf(gavaged.ps)

#generate color objects
color_objs.akk_free <- create_color_dfs(mdf.akk_free,
                                        selected_groups = c("Actinobacteriota", "Bacteroidota", "Firmicutes"),
                                        cvd = TRUE)
color_objs.akk_free <- reorder_samples_by(color_objs.akk_free$mdf, color_objs.akk_free$cdf)

color_objs.gavaged <- create_color_dfs(mdf.gavaged,
                                        selected_groups = c("Actinobacteriota", "Bacteroidota", "Firmicutes", "Verrucomicrobiota"),
                                        cvd = TRUE)
color_objs.gavaged <- reorder_samples_by(color_objs.gavaged$mdf, color_objs.gavaged$cdf)

#Extract
mdf.akk_free <- color_objs.akk_free$mdf
cdf.akk_free <- color_objs.akk_free$cdf

mdf.gavaged <- color_objs.gavaged$mdf
cdf.gavaged <- color_objs.gavaged$cdf
```

```{r}
akk_free.legend <- custom_legend(mdf.akk_free, cdf.akk_free, legend_text_size = 15)

akk_free.plot <- plot_microshades(mdf.akk_free, cdf.akk_free, x = "SampleName") +
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 8)) +
  facet_wrap(~Group, scales = "free_x") +
  theme(strip.text.x = element_text(size = 8))

plot_grid(akk_free.plot, akk_free.legend, rel_widths = c(1, .25))
```

```{r}
gavaged.legend <- custom_legend(mdf.gavaged, cdf.gavaged, legend_text_size = 15)

gavaged.plot <- plot_microshades(mdf.gavaged, cdf.gavaged, x = "Day") +
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 8)) +
  scale_x_discrete(limits=c(3, 7, 14, 21, 28)) +
  facet_wrap(~Group, scales = "free_x") +
  theme(strip.text.x = element_text(size = 8))

plot_grid(gavaged.plot, gavaged.legend, rel_widths = c(1, .25))
```

Visually, there appear to be some differences between the Akk-free/late PBS mice and the akk-gavaged mice. Most striking is that Bifidobacterium are only present in the akk-free and late PBS mice. There also appears to be a greater abundance of Dubosiella in the akkermansia-gavaged mice.

#Differential Abundance
LEfSe

We want to focus on the endpoint samples.
```{r}
genus.ps <- gavaged.ps %>%
  tax_glom(taxrank = "Genus")

lef <- lefse(genus.ps,
             class = "Group",
             multicls_strat = TRUE)
```

```{r}
plot_ef_bar(lef)
```

It does appear that we were on to something thinking that Bifidobacterium were more indicative of mice unexposed to akkermansia. Akkermanisa are indicative of WT-gavaged mice which makes sense. Ruminococcus are indicative of 2086-gavaged mice and Blautia are also indicative of PBS mice, neither of which showed up as being a part of the top 20 taxa overall.

#Akkermansia
```{r}
akks.ps <- rel.ps %>%
  subset_taxa(Genus == "Akkermansia") %>%
  tax_glom(taxrank = "Genus")
print(akks.ps)
```

```{r}
akks <- akks.ps %>%
  otu_table() %>%
  as.data.frame()
```

Beta
```{r}
even.ps <- transform_sample_counts(just.dustin.ps, function(x) 1E6 * x/sum(x))
```

```{r}
even.nmds_bs <- ordinate(even.ps, "NMDS", "bray")
```

```{r}
plot_ordination(even.ps, even.nmds_bs, color = "Day", shape = "Group") +
  geom_point(size = 5)
```

#SFB
```{r}
clostridiaceae.ps <- rel.ps %>%
  subset_taxa(Family == "Clostridiaceae")
print(clostridiaceae.ps)
```

```{r}
clostridiaceae <- clostridiaceae.ps %>%
  otu_table() %>%
  as.data.frame()
```

```{r}
write.csv(clostridiaceae, "clostridiaceae_asvs.txt", row.names = FALSE)
```


#Ruminococcus
```{r}
rumino.ps <- rel.ps %>%
  subset_taxa(Genus == "Ruminococcus") %>%
  tax_glom(taxrank = "Genus")
print(akks.ps)
```

```{r}
rumino <- rumino.ps %>%
  otu_table() %>%
  as.data.frame()
```