---
title: "Emi's Analysis"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(microbiomeMarker)
library(microshades)
library(speedyseq)
library(cowplot)
```

#Define Paths
```{r}
rds <- "/work/kdm65/2021-Dec.rds"
```

Load and check phyloseq object
```{r}
all_mice.ps <- readRDS(rds)
print(all_mice.ps)
```

```{r}
sample_variables(all_mice.ps)
```

Subset for just Dustin's mice and remove bad samples
```{r}
just.emi.ps <- all_mice.ps %>%
  subset_samples(Owner == "Emi") %>%
  subset_samples(!SampleName %in% c("1","12", "13"))
just.emi.ps
```

```{r}
plot_bar(just.emi.ps)
```
None of these samples stand out as having super low read counts, so I don't need to prune any out.

#Relative Abundance
```{r}
rel.ps <- just.emi.ps %>%
  transform_sample_counts(function(x) x/sum(x))
```

#Abundance Plots
```{r}
AbundancePlot = function(ps, tax_level) {
  ps %>%
    plot_bar(fill = tax_level, x = "Age") +
    labs(y = "Relative Abundance") +
    geom_bar(aes_string(color = tax_level, fill = tax_level),
             stat = "identity",
             position = "stack") +
    scale_x_discrete(limits = c(76, 98, 118, 138))
}
```

```{r}
females.ps <- rel.ps %>%
  subset_samples(Sex == "f")
males.ps <- rel.ps %>%
  subset_samples(Sex == "m")
```

```{r}
legend <- cowplot::get_legend(AbundancePlot(females.ps, "Phylum"))

female.plot <- AbundancePlot(females.ps, "Phylum") +
  facet_grid(~Mouse_SID) +
  theme(legend.position = "none") +
  labs(title = "Females")
male.plot <- AbundancePlot(males.ps, "Phylum") +
  facet_grid(~Mouse_SID) +
  theme(legend.position = "none") +
  labs(title = "Males")

sexes <- plot_grid(female.plot, male.plot, ncol = 1)
plot_grid(sexes, legend, rel_widths = c(3,1))
```

#Pruning
Prune to the top 20 genera
```{r}
rel.ps %>%
  taxa_sums %>%
  sort(decreasing = TRUE) %>%
  names %>%
  head(20) ->
  top20

rel.ps %>%
  prune_taxa(top20, .) ->
  prune20.rel
prune20.rel
```

```{r}
pruned.female.ps <- prune20.rel %>%
  subset_samples(Sex == "f")

pruned.male.ps <- prune20.rel %>%
  subset_samples(Sex == "m")
```

```{r}
pruned.legend <- cowplot::get_legend(AbundancePlot(pruned.female.ps, "Genus"))

pruned.female.plot <- AbundancePlot(pruned.female.ps, "Genus") +
  facet_grid(~Mouse_SID) +
  theme(legend.position = "none") +
  labs(title = "Females")
pruned.male.plot <- AbundancePlot(pruned.male.ps, "Genus") +
  facet_grid(~Mouse_SID) +
  theme(legend.position = "none") +
  labs(title = "Males")

pruned.sexes <- plot_grid(pruned.female.plot, pruned.male.plot, ncol = 1)
plot_grid(pruned.sexes, pruned.legend, rel_widths = c(3,1))
```

Trying out a colorblind friendly plotting package

```{r}
#melt to dataframes
mdf.female <- prep_mdf(pruned.female.ps)
mdf.male <- prep_mdf(pruned.male.ps)

#generate color objects
color_objs.female <- create_color_dfs(mdf.female,
                                        selected_groups = c("Bacteroidota", "Firmicutes", "Verrucomicrobiota"),
                                        cvd = TRUE)
color_objs.female <- reorder_samples_by(color_objs.female$mdf, color_objs.female$cdf)

color_objs.male <- create_color_dfs(mdf.male,
                                        selected_groups = c("Firmicutes", "Verrucomicrobiota"),
                                        cvd = TRUE)
color_objs.male <- reorder_samples_by(color_objs.male$mdf, color_objs.male$cdf)

#Extract
mdf.female <- color_objs.female$mdf
cdf.female <- color_objs.female$cdf

mdf.male <- color_objs.male$mdf
cdf.male <- color_objs.male$cdf
```

```{r}
cb.female.legend <- custom_legend(mdf.female, cdf.female, legend_text_size = 15)

cb.female.plot <- plot_microshades(mdf.female, cdf.female, x = "Age") +
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 8)) +
  scale_x_discrete(limits = c(76, 98, 118, 138)) +
  facet_wrap(~Mouse_SID) +
  theme(strip.text.x = element_text(size = 8))

#plot_grid(cb.female.plot, cb.female.legend, rel_widths = c(1, .25))
```

```{r}
cb.male.legend <- custom_legend(mdf.male, cdf.male, legend_text_size = 15)

cb.male.plot <- plot_microshades(mdf.male, cdf.male, x = "Age") +
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 8)) +
  scale_x_discrete(limits = c(76, 98, 118, 138)) +
  facet_wrap(~Mouse_SID) +
  theme(strip.text.x = element_text(size = 8))

#plot_grid(cb.male.plot, cb.male.legend, rel_widths = c(1, .25))
```

```{r}
cb.sexes <- plot_grid(cb.female.plot, cb.male.plot, ncol = 1)
plot_grid(cb.sexes, cb.female.legend, rel_widths = c(3,1))
```

#Differential Abundance
LEfSe
```{r}
genus.ps <- rel.ps %>%
  tax_glom(taxrank = "Genus")

lef <- lefse(genus.ps,
             class = "Sex",
             multicls_strat = TRUE)
```

Filter for low abundance taxa?
```{r}
mean_cutoff = 0.001
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

pct0.1_prune.rel <- rel.ps %>%
  filter_taxa(mean_test, prune = TRUE)
```

```{r}
genus.ps <- pct0.1_prune.rel %>%
  tax_glom(taxrank = "Genus")

lef <- lefse(genus.ps,
             class = "Sex",
             multicls_strat = TRUE)
```

```{r}
plot_ef_bar(lef)
```

#Akkermansia
```{r}
akks.ps <- rel.ps %>%
  subset_taxa(Genus == "Akkermansia") %>%
  tax_glom(taxrank = "Genus")
print(akks.ps)
```

```{r}
akks <- akks.ps %>%
  otu_table() %>%
  as.data.frame()
```